Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-12 paper.tex
wording

\sklearn is explicitly a library,
i.e. a collection of classes (in the object-oriented sense)
that users ``import'' into Python scripts/programs.
Use of the package requires basic Python programming knowledge,
and no command-line interface, let alone a GUI,
is offered for non-programmer users.
Instead, interactive use is made possible
by the Python interactive interpreter,
and its enhanced replacement IPython \citep{perez2007ipython},
which offers a \textsc{matlab}-like working environment
specifically designed for scientific use.

From its core, the library has been designed to tie in with the set
of numeric and scientific packages centered around the NumPy and SciPy libraries.
Of these, NumPy \citep{vanderwalt2011} augments Python
with a numeric array datatype and associated array computing primitives,
while SciPy \citep{varoquaux2013scipy} extends it further
by making common numeric operations available,
either by implementing those in Python/NumPy,
or by wrapping existing C/C{}\verb!++!/Fortran implementations.
In a similar vein, \sklearn extends the general-purpose language Python
with a set of composable machine learning operations.

Started in 2009, \sklearn is developed by an international team
of a dozen of core developers, mostly researchers from various fields
(e.g., machine learning, natural language processing, or neuroscience).
The project also greatly benefits from many occasional
contributors proposing small code or documentation fixes or improvements.
Development proceeds on GitHub (\url{https://github.com/scikit-learn}),
a platform which greatly facilitates this kind of collaboration.
Because of the large number of developers,
particular emphasis is put on keeping the project maintainable.
In particular, code must follow specific
quality guidelines, such as style consistency and unit-test coverage. In
addition, complete documentation and examples are required for all
features.
Finally, additions to the code base must pass code review
by at least two developers not involved in the implementation
of the proposed addition.

Within the past few years, the project has proved valuable to an ever increasing
user base. Its popularity can be gaged from various indicators such as hundreds
of citations in scientific publications of a previous paper
\citep{pedregosa2011}, successes in various machine learning challenges
(\emph{e.g.} Kaggle), or significant statistics derived from our repositories
and mailing lists.  At the time of writing, the project is watched by 1309
people and forked 667 time on GitHub, and the mailing-list receives 300 mails
per month.
The version control logs for the development version of scikit-learn
% ddaa494c116e3c16bf032003c5cccbed851733d2
show 183 contributors to the project's source code and documentation
(after manual deduplication).

>>>>

The project is a programming library, i.e. a collection of classes (in the
object-oriented sense) that users import into Python programs. Its use
therefore requires basic Python programming knowledge. No command-line
interface, let alone a GUI, is offered for non-programmer users. Instead,
interactive use is made possible by the Python interactive interpreter, and its
enhanced replacement IPython \citep{perez2007ipython}, which offers a
\textsc{matlab}-like working environment specifically designed for scientific
use.

From its core, the library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a numeric array datatype and
associated array computing primitives while SciPy \citep{varoquaux2013scipy}
extends it further by making common numeric operations available, either by
implementing those in Python/NumPy or by wrapping existing
C/C{}\verb!++!/Fortran implementations. In a similar vein, \sklearn extends
Python with a set of composable machine learning operations.

Started in 2009, \sklearn is developed by an international team of a dozen of
core developers, mostly researchers from various fields (e.g., machine learning,
natural language processing, or neuroscience). The project also greatly benefits
from many occasional contributors proposing small code or documentation fixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
In addition, complete documentation and examples are required for all features.
Finally, additions to the code base must pass code review by at least two
developers not involved in the implementation of the proposed addition.

Within the past few years, the project has proved valuable to an ever increasing
user base. Its popularity can be gaged from various indicators such as hundreds
of citations in scientific publications, successes in various machine learning challenges
(\emph{e.g.} Kaggle), or significant statistics derived from our repositories
and mailing lists.  At the time of writing, the project is watched by 1309
people and forked 667 time on GitHub, and the mailing-list receives 300 mails
per month.
The version control logs for the development version of scikit-learn
% ddaa494c116e3c16bf032003c5cccbed851733d2
show 183 contributors to the project's source code and documentation
(after manual deduplication).
